---
title: "DNA_data"
output: html_document
date: "2023-03-02"
---

Load libraries 
```{r message=FALSE, warning=FALSE}
library(readxl)
library(mapview)
library(sf)
library(magrittr)
library(stringr)
library(dplyr)
library(janitor)
```


# Data

## eDNA
```{r}
setwd('C:/Users/greco/OneDrive - USherbrooke/Maitrise/Projet de maitrise/data/ADN/raw')

```

### Chatauguay

Open data for the Chatauguay river. The third sheet will be used now since it represents the presence and absence of species without the number of DNA sequences obtained. 
```{r}
chatauguay_data<-read_excel('C:/Users/greco/OneDrive - USherbrooke/Maitrise/Projet de maitrise/data/ADN/raw/Résultat_ADNe_Chatauguay_2021_.xlsx',sheet='présence_absence')
chatauguay_df<-as.data.frame(chatauguay_data) 
```

Before building the community matrix we will plot where the data was taken. To do so, we remove the presence absence data to keep only coordinates and environmental variable. 

Subset:
```{r}
chatauguay_env<-chatauguay_df[c(3:60),c(1:9)]
```

Make it a spatial object and plot it on a map.
```{r}
chatauguay_sf <- st_as_sf(x =chatauguay_env, 
                        coords = c("Longitude","Latitude"),
                        crs = "+proj=longlat +ellps=WGS84 +no_defs")

mapview(chatauguay_sf)
```

### Saint-François

Same steps as in Chatauguay River.
```{r}
saint_francois_data<-read_excel('C:/Users/greco/OneDrive - USherbrooke/Maitrise/Projet de maitrise/data/ADN/raw/Résultat_ADNe_Saint-François_2021_.xlsx',sheet='présence_absence')
saint_francois_df<-as.data.frame(saint_francois_data)

#subset for only environmental data with coordinates
saint_francois_env<-saint_francois_df[c(3:146),c(1:9)]

#make it a spatial object
saint_francois_sf <- st_as_sf(x =saint_francois_env, 
                          coords = c("Longitude","Latitude"),
                          crs = "+proj=longlat +ellps=WGS84 +no_defs")

mapview(saint_francois_sf)+chatauguay_sf

```

## Dams

```{r}

dam_raw <-
  read_excel(
    'C:/Users/greco/OneDrive - USherbrooke/Maitrise/Projet de maitrise/data/environment/repertoire_des_barrages.xls'
  )

```

The head does not contain important column names, so I will set the first row in the data frame as the header. 
```{r}
dam<-janitor::row_to_names(dam_raw, 1, remove_rows_above = FALSE) 
dam<-as.data.frame(dam) #turn into df
```
I want to make it a spatial object but one dam does not contain latitude ond longitude data. I decided to remove this dam of the dataset. 
```{r}
dam_na<-dam[!is.na(dam$`Latitude (NAD 83)`),]
```
With the NA removed I can now make it a spatial object and plot it on a map. 
```{r}
dam_sf<- st_as_sf(x =dam_na, 
                          coords = c("Longitude (NAD 83)","Latitude (NAD 83)"),
                          crs = "+init=epsg:4269 +proj=longlat +ellps=GRS80 +datum=NAD83
+no_defs +towgs84=0,0,0
") 

mapview(dam_sf)
```

This data set represents dams in the whole province. I will use the zones delinieated by the government of Quebec to select the dams that are present in the zone where our rivers can be found. The shapefile of these zones can be found here : https://www.donneesquebec.ca/recherche/dataset/grhq . The whole hydrography can be downloaded here: https://mrnf.gouv.qc.ca/repertoire-geographique/reseau-hydrographique-grhq/ . 

```{r}
hydro_cut<-st_read('C:/Users/greco/OneDrive - USherbrooke/Maitrise/Projet de maitrise/data/environment/decoupage')

mapview(list(hydro_cut,saint_francois_sf,chatauguay_sf),col.regions=list('red','blue','blue'))

```
All the sampling was done in zone 03 which represents ID 1. I will then subset to only have a polygon of zone 3.
```{r}
hydro_03<-hydro_cut[hydro_cut$Bloc=='03',]

```




